Hierarchical Controller Architecture in Software Defined Networks

### 2013-12-16 paper/paper.tex
minor changes

\section{Introduction}
%Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature. Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements. Some classes of control applications are routing, traffic engineering (TE), enforcing network policies, etc. 
In traditional networks, there is a tight coupling and a one to one mapping between control elements and forwarding elements, with one pair on each switch. Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules. It is the control application which runs distributed protocols to synchronize the state of forwarding elements. The design of these distributed protocols differs from application to application. For example, link state routing makes switches share their neighbors while distance vector routing makes switches share their routing tables.

@@

Conventional control plane architectures involve multiple peer controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application. We referred to such an architecture as \emph{flat controller architecture}. Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT). Consequently, such an architecture results in huge communication overhead between controllers, which may hinders the scalability of the system. Besides, network state inconsistency is quite common in flat architecture which degrades the application performance significantly \cite{logically-centralized}.

The state inconsistency and huge communication overhead problem arises from the fact that each controller tries to maintain the complete global view of the network. Motivated by this observation, we propose a hierarchical controller architecture, in which controllers are arranged in a tree structure and exposes a switch abstraction to it's parent. We show that hierarchical architecture significantly reduces the communication overhead between controllers while modestly compromising on optimality.

>>>>

\section{Introduction}
%Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature. Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements. Some classes of control applications are routing, traffic engineering (TE), enforcing network policies, etc. 
In traditional networks, there is a tight coupling and a one-to-one mapping between control elements and forwarding elements, with one pair on each switch. Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules. It is the control application which runs distributed protocols to synchronize the state of forwarding elements. The design of these distributed protocols differs from application to application. For example, link state routing makes switches share their neighbors while distance vector routing makes switches share their routing tables.

@@

Conventional control plane architectures involve multiple peer controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application. We refer to such an architecture as \emph{flat controller architecture}. Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT). Consequently, such an architecture results in huge communication overhead between controllers, which may hinders the scalability of the system. Besides, network state inconsistency is quite common in flat architecture which degrades the application performance significantly \cite{logically-centralized}.

The state inconsistency and huge communication overhead problem arises from the fact that each controller tries to maintain the complete global view of the network. Motivated by this observation, we propose a hierarchical controller architecture, in which controllers are arranged in a tree structure and exposes a switch abstraction to its parent. We show that hierarchical architecture significantly reduces the communication overhead between controllers while modestly compromising on optimality.
